


\section{Leakage-Safe Retrieval Teacher Construction}
\label{appx:Offline_Teacher_Construction_Algorithm}

\input{tables/Offline-Teacher-Construction}


\section{Theoretical Analysis}
\label{sec:theory}

TS-Memory aims to capture the \emph{adaptivity} of retrieval-based forecasting without paying retrieval-time
latency at deployment. This appendix provides intuition from three complementary views:
\begin{enumerate}
    \item \textbf{Privileged distributional supervision.} Offline retrieval induces a discrete,
    context-conditioned predictive distribution. The teacher quantiles $\widetilde{\mathbf{Q}}_t$
    are weighted-quantile solutions of a pinball objective, linking our targets to local
    non-parametric conditional quantile estimation (Section~\ref{subsec:teacher}).
    \item \textbf{Shift-aligned retrieval.} The alignment in Eq.~(\ref{eq:tsm-shift})--(\ref{eq:tsm-align})
    and reranking explicitly handle additive level shifts, so neighbors are compared mainly by
    \emph{shape} rather than absolute scale (Section~\ref{subsec:shift}).
    \item \textbf{Conservative distillation and deployment.} Reliability-aware gating reduces negative
    transfer by down-weighting (or disabling) distillation when retrieval is unreliable, and
    deployment-time quantile fusion is a convex combination with an immediate pinball-risk guarantee
    (Section~\ref{subsec:reliability}--\ref{subsec:deployment}).
\end{enumerate}

\subsection{Offline Retrieval Supervision}
\label{subsec:teacher}

Given a query context $\mathbf{X}_t$, let $\{\mathbf{Y}^{(k)}_{\text{align}}\}_{k=1}^{K}$ be the top-$K$
aligned future windows retrieved from the \emph{training-only} index, with weights
$\{w_k\}_{k=1}^{K}$ such that $\sum_{k=1}^{K} w_k = 1$ (Eq.~\ref{eq:tsm-ret-weights}).
This defines a teacher available only during offline training; the memory module learns to reproduce its
distributional information without test-time search.
For each horizon--variable index $u\in\mathcal{U}$, define samples
$v_k \triangleq Y^{(k)}_{\text{align},u}$ and the weighted empirical measure
\begin{equation}
\widehat{\mathcal{P}}_{t,u}
\triangleq
\sum_{k=1}^{K} w_k \,\delta_{v_k},
\end{equation}
which is a discrete predictive distribution supported on retrieved futures. It naturally captures uncertainty
(and even multi-modality) from diverse neighbors, providing richer supervision than a single point label.

\noindent\textbf{Weighted quantiles as pinball minimizers.}
For any quantile level $q\in(0,1)$, consider the weighted pinball risk
\begin{equation}
R_{t,u}(z;q)
\triangleq
\mathbb{E}_{V\sim \widehat{\mathcal{P}}_{t,u}}
\big[\rho_{q}(z,V)\big]
=
\sum_{k=1}^{K} w_k \,\rho_{q}(z,v_k),
\end{equation}
where $\rho_q(z,v)=(v-z)\big(q-\mathbb{I}_{v<z}\big)$.
By a standard subgradient argument, any minimizer $z^\star$ satisfies
\begin{equation}
\sum_{k: v_k < z^\star} w_k \le q \le \sum_{k: v_k \le z^\star} w_k.
\label{eq:weighted-quantile-condition}
\end{equation}
If $\widehat{\mathcal{P}}_{t,u}$ has atoms, the minimizer can be non-unique; Eq.~(\ref{eq:tsm-teacher-quantile})
chooses the \emph{left} (lower) weighted quantile satisfying Eq.~(\ref{eq:weighted-quantile-condition}).
Thus, distilling $\widetilde{\mathbf{Q}}_t$ into $g_{\phi}(\mathbf{X}_t)$ can be viewed as \emph{amortizing} a
local non-parametric conditional quantile estimator $\{(v_k,w_k)\}_{k=1}^{K}$ into a constant-time module.

\noindent\textbf{Kernel view and distributional metrics.}
The teacher can also be interpreted as a kernel-smoothed local estimate of the conditional predictive
distribution around $\mathbf{X}_t$ in the frozen embedding space, since
$w_k \propto \exp(-\psi(d_k)/\tau_{\text{ret}})$.
Moreover, many distributional scores can be expressed (exactly or approximately) as integrals of pinball
losses over quantile levels (e.g., CRPS for univariate targets). Hence, improving quantile alignment across
$\{q_j\}$ tends to improve distributional evaluation.

\subsection{Shift Alignment and Reranking}
\label{subsec:shift}

A common non-stationarity is a channel-wise level offset. We model each window as
\[
\mathbf{X}_t = \mathbf{Z}_t + \mathbf{b}_t,
\qquad
\mathbf{Y}_t = \mathbf{W}_t + \mathbf{b}_t,
\]
where $\mathbf{Z}_t,\mathbf{W}_t$ are shift-free patterns and $\mathbf{b}_t\in\mathbb{R}^{C}$ is a per-window
offset (constant within the context and forecast horizon).

For a retrieved candidate $i$ with offset $\mathbf{b}_i$, the trailing-mean shift in Eq.~(\ref{eq:tsm-shift})
estimates $\mathbf{b}_t-\mathbf{b}_i$:
\[
\boldsymbol{s}_i
=
\mathrm{mean}\!\left(\mathbf{X}_{t,L-m+1:L}\right)
-
\mathrm{mean}\!\left(\mathbf{X}^{(i)}_{L-m+1:L}\right)
\approx
\mathbf{b}_t-\mathbf{b}_i.
\]
After applying Eq.~(\ref{eq:tsm-align}), we obtain approximately
\begin{equation}
\mathbf{X}^{(i)}_{\text{align}} \approx \mathbf{Z}^{(i)} + \mathbf{b}_t,
\qquad
\mathbf{Y}^{(i)}_{\text{align}} \approx \mathbf{W}^{(i)} + \mathbf{b}_t,
\end{equation}
so retrieved futures are expressed in the same offset frame as the query.

\noindent\textbf{Why reranking matters.}
TS-Memory reranks by $\text{score}_i=\|\mathbf{X}_t-\mathbf{X}^{(i)}_{\text{align}}\|_1$.
Under the additive-shift model,
\[
\mathbf{X}_t-\mathbf{X}^{(i)}_{\text{align}}
\approx
\mathbf{Z}_t-\mathbf{Z}^{(i)},
\]
so reranking compares \emph{shift-free shapes} rather than absolute levels, reducing scale-mismatch artifacts.
This complements Instance Normalization in Eq.~(\ref{eq:tsm-inorm}): normalization encourages shift-invariant
learning, while alignment (and inverse normalization) restores outputs to the correct absolute scale.

\subsection{Reliability-Aware Distillation}
\label{subsec:reliability}

Retrieval-based distillation can hurt when neighbors are noisy or mismatched. TS-Memory reduces this risk
with a two-stage reliability mechanism: (i) a confidence proxy from retrieval separation, and (ii) an
advantage test that activates distillation only when the teacher is better than the frozen backbone.

\noindent\textbf{Confidence as retrieval separation.}
We use $\mathrm{Conf}_t=\max_k w_k$ (Eq.~\ref{eq:tsm-conf}). Let $d_{(1)}\le d_{(2)}\le\cdots\le d_{(K)}$ be
sorted distances and $\Delta=d_{(2)}-d_{(1)}$ the top-1 margin. Since the weights are a softmax over distances,
\begin{equation}
\mathrm{Conf}_t = w_{(1)}
\ge \frac{1}{1+(K-1)\exp(-\Delta/\tau_{\text{ret}})}.
\label{eq:conf-margin}
\end{equation}
Larger $\mathrm{Conf}_t$ indicates a clearer nearest-neighbor match.

\noindent\textbf{Distill only when beneficial.}
Confidence alone can be misleading, so TS-Memory further gates distillation using an advantage test
(Eq.~\ref{equ:err}):
\begin{equation}
\chi_t
=
\mathbb{I}\!\left(
\mathrm{err}^{T}_t + \epsilon_{\text{gate}}
<
\mathrm{err}^{\text{base}}_t
\right),
\qquad
\omega_t
=
\chi_t \cdot \mathrm{Conf}_t^{\gamma}.
\end{equation}
The alignment loss (Eq.~\ref{eq:tsm-loss-align}) is weighted by $\omega_t$; when $\chi_t=0$, distillation
contributes no gradient, avoiding systematic negative transfer.

\noindent\textbf{Residual distillation and stabilization.}
Besides matching teacher quantiles (Eq.~\ref{eq:tsm-dq}), TS-Memory distills an incremental median correction
over the frozen backbone (Eq.~\ref{eq:tsm-delta}--\ref{eq:tsm-ddelta}):
\[
\Delta^{\text{mem}}_{t,u}
=
\widehat{Q}^{\text{mem}}_{t,j^\star,u}
-
\widehat{Q}^{\text{base}}_{t,j^\star,u},
\qquad
\Delta^{T}_{t,u}
=
\widetilde{Q}_{t,j^\star,u}
-
\widehat{Q}^{\text{base}}_{t,j^\star,u}.
\]
This focuses learning on the teacher's \emph{improvement} over a strong frozen prior. When retrieval is
uncertain, the anchoring loss (Eq.~\ref{eq:tsm-loss-anchor}) pulls the memory median toward the backbone
median with weight $(1-\omega_t)$, and the quantile crossing penalty (Eq.~\ref{eq:tsm-loss-cross}) encourages
monotonicity across quantile levels.

\subsection{Deployment via Convex Quantile Fusion}
\label{subsec:deployment}

At inference time, TS-Memory performs no retrieval and fuses backbone and memory forecasts by quantile-wise
interpolation (Eq.~\ref{eq:tsm-fuse}):
\begin{equation}
\widehat{\mathbf{Q}}^{\text{final}}_t
=
(1-\alpha)\widehat{\mathbf{Q}}^{\text{base}}_t
+
\alpha \widehat{\mathbf{Q}}^{\text{mem}}_t,
\qquad
\alpha\in[0,1].
\end{equation}
We tune a single $\alpha$ on validation to preserve quantile coherence and keep tuning 1D.

Because $\rho_q(z,y)$ is convex in $z$, for any fixed $(t,j,u)$ and target $y$,
\begin{equation}
\begin{aligned}
\rho_{q_j}\!\Big(
(1-\alpha)\widehat{Q}^{\text{base}}_{t,j,u}
+
\alpha \widehat{Q}^{\text{mem}}_{t,j,u},
\, y
\Big)
\le\;&
(1-\alpha)\rho_{q_j}\!\Big(\widehat{Q}^{\text{base}}_{t,j,u},y\Big) \\
&+
\alpha \rho_{q_j}\!\Big(\widehat{Q}^{\text{mem}}_{t,j,u},y\Big).
\end{aligned}
\end{equation}
Averaging over horizons, variables, and quantiles yields that the fused empirical pinball risk is
upper-bounded by the same convex combination of the individual risks. Therefore, the best $\alpha\in[0,1]$
(on any evaluation set) is never worse than choosing either endpoint predictor. In practice, since the
objective is convex (piecewise-linear) in $\alpha$, a small grid search on validation pinball/CRPS is enough.
Since the risk is convex in $\alpha$, any local optimum is global; increasing grid resolution only refines the minimizer.
If both quantile sets are monotone in $q_j$, their convex combination remains monotone, so fusion preserves
distribution validity.

\noindent\textbf{Extension to point forecasting.}
When point forecasts are needed, we extract a point estimate from quantiles. Let $q_{j^\star}$ be closest to
$0.5$. If $0.5\in\mathcal{Q}$, use the median; otherwise interpolate between $q_{j_a}<0.5<q_{j_b}$:
\begin{equation}
\widehat{p}_{t,u}
=
\widehat{Q}_{t,j_a,u}
+
\frac{0.5-q_{j_a}}{q_{j_b}-q_{j_a}}
\Big(
\widehat{Q}_{t,j_b,u}
-
\widehat{Q}_{t,j_a,u}
\Big).
\end{equation}
Point forecast fusion is analogous:
\begin{equation}
\widehat{p}^{\text{final}}_{t,u}
=
(1-\alpha)\widehat{p}^{\text{base}}_{t,u}
+
\alpha \widehat{p}^{\text{mem}}_{t,u},
\end{equation}
with $\alpha$ tuned on validation MAE/MSE. For point-only backbones, treat the backbone output as
$\widehat{p}^{\text{base}}$.


