\section{Future Work}
\label{sec:future_work}
TS-Memory motivates a broader \emph{retrieval-to-memory} paradigm, where retrieval-induced distributional knowledge is compiled into a compact module for retrieval-free deployment. Future work can extend this paradigm along several directions.

\noindent\textbf{Continual and streaming memory updates.}
An appealing direction is enabling memory modules to evolve with newly observed data in non-stationary environments, while preserving retrieval-free inference.
This includes rolling-buffer teacher refresh, lightweight incremental distillation, and continual-learning regularizers that stabilize memory representations across drift.

\noindent\textbf{Richer teachers and reliability learning.}
Future work may explore stronger offline teacher signals, such as hybrid similarity metrics that combine embedding distance with correlation- or frequency-aware matching, as well as teacher ensembles across multiple retrieval spaces.
In addition, learned reliability predictors calibrated to downstream utility can provide a principled way to decide when teacher signals should guide memory learning.

\noindent\textbf{Broader alignment families for distribution shift.}
Beyond basic alignment settings, it is promising to study leakage-safe alignment families that capture a wider spectrum of non-stationarities, including scale changes, trend deformations, and temporal warping.
Such alignment can improve neighbor correspondence and yield sharper, more informative teacher distributions across domains.

\noindent\textbf{Input-adaptive fusion and uncertainty-aware calibration.}
Another direction is to make fusion between backbone and memory predictions context-conditioned, allowing the model to adapt memory reliance based on uncertainty and domain mismatch.
This could further improve probabilistic calibration by coupling fusion decisions with uncertainty-aware confidence modeling.

\noindent\textbf{Scalable teacher compilation and broader tasks.}
To support large-scale and multi-domain deployment, future work can develop more efficient teacher compilation pipelines, including approximate neighbor search, teacher-set distillation, and shared teacher banks across related domains.
Beyond long-horizon forecasting, retrieval-to-memory distillation may extend naturally to imputation, anomaly detection, and decision-focused forecasting, where retrieved analogs can provide privileged distributional supervision.






\input{tables/different-TSFMs-with-TS-Memory}

\input{tables/ragvsTS-Memory}

\input{tables/efficiency_query}

\input{tables/efficiency_time}


\input{tables/different-size-of-Chronos}


\input{tables/plugmem_merged_table}

\input{tables/TSMemoryLoRA-PlugMem-Size}