\section{Related Works}
\subsection{LLMs for Time Series Analysis}
Previous studies mainly focus on utilizing Large Language Models in conventional time series analytical tasks such as forecasting and anomaly detection. Their motivation primarily lies in endowing time series models with generalization capabilities transfered from LLMs, which can help alleviate the domain-specific limitations that arise when large-scale time series corpus is scarce. Among them, TimeLLM~\citep{timellm}, TimeVLM~\citep{timevlm}, ChatTime~\citep{wang2025chattime}, UniTime~\citep{liu2024unitime}, and CALF~\citep{liu2025calf} demonstrate certain performance. However, some studies~\citep{zhou2024can,tan2024language} point out that LLMs do not achieve significant performance compared to deep time series models but possess larger overheads. 

On the other hand, Large Language Models are suitable for tasks with high requirements for cross-modality and interpretability. Therefore, recent studies like ChatTS~\citep{xie2024chatts}, TimeOmni-1~\citep{guan2025timeomni1incentivizingcomplexreasoning}, and ITformer~\citep{leiitformer} pay attention to enhance the inherent reasoning capability of LLMs in TSQA tasks, helping data scientists solve problems from vertical domains such as finance, weather, and aircraft engine~\citep{leiitformer}. With the rise of exploration in this field, some TSQA benchmarks are proposed to fairly evaluate the performance of TSRMs. Among them, MTBench~\citep{chen2025mtbench}, TimeSeriesExam~\citep{cai2024timeseriesexam}, and TimeSeriesGym~\citep{cai2025timeseriesgym} gain attention. However, most of efforts are limited to improving the ability of LLM themselves to handle TSQA tasks, neglecting that numerical ability is a weakness of LLMs. In this work, we introduce TimeART to pionner the exploration of agentic time series reasoning, combining the advantages of analytical tools in numerical processing with the strong reasoning capability of LLMs. 

\subsection{Agents for Time Series Analysis}
Recent studies on time series agents mainly focus on some conventional tasks, such as representation learning~\citep{zhou2025merit}, classification~\citep{sui2025training}, and AutoML for forecasting~\citep{zhao2025timeseriesscientist,wang2025aries}. These studies leverage the decision-making capability of LLM-based agent or multi-agents to pursue automation or few-shot inference, so that they can replace repetitive labor to some extent. However, these tasks are already well solved by some conventional methods. Though some studies focus on time series reasoning agents~\citep{liu2025tsagenttimeseriesreasoning,ye2025beyond}, they put emphasis on designing a fixed procedure or do not fully develop the diversity of external analytical tools, and not well adapt LLMs to the time series domain through training. In this work, we incorporate TimeART with strong tools, and propose TimeToolBench to finetune TSRMs with a novel training strategy. 
